\section{Probabilistic Methods for Ordinary Differential Equations}

The numerical integration of Ordinary Differential Equations (ODE's) is a topic of numerical mathematics which has been deeply studied in the last decades. The main focus of researchers was providing and analyzing methods endowed with properties such as convergence, conservation of invariants and stability. One class of methods which has proved to be particularly successful are the Runge-Kutta methods, which can alternatively provide some of the properties above. It is well-known that these methods are based on the choice of a discretization grid identified by a time step, that we denote by $h$, where the numerical solution will be computed. The introduction of a temporal discretization results in an error, which will be proportional to the time step $h$ at a power $q$. The quantity $q$ is referred to as the \textit{order} of the numerical method, and gives an indication of the convergence properties of the numerical solution towards the exact solution of the ODE. \\
Recently, researchers have been interested in providing probabilistic numerical solutions to ODE's \cite{CGS16}, thus introducing probabilistic measures which describe the numerical solution instead of punctual solutions. In this way it is possible to provide a quantification of the numerical error introduced at final time, which will depend both on the chosen discretization step and on the properties of the considered equation. \\
Let us consider $f\colon \R^d \to \R^d$ and the following ODE
\begin{equation}\label{eq:ODE}
\begin{aligned}
	u'(t) &= f(u), &&  t \in (0, T], \\
	u(0)  &= u_0, && u_0 \in \R^d.
\end{aligned}
\end{equation}
Given a time step $h >0$, we now consider a discretization of the time interval $[0, T]$ defined as $t_k = kh$, with $k = 0, \ldots, N$, and with $T = Nh$. Let us now consider the numerical solution numerical solution $U_k$ at time $t_k$ given by a Runge-Kutta methods. We can write one step of the time integration as 
\begin{equation}\label{numericalODE}
\begin{aligned}
	U_{k+1} &= \Psi(U_k), \quad k = 0, \ldots, N-1, \\
	U_{0} &= u_0, 
\end{aligned}
\end{equation}
where $\Psi$ defines the flow map of a Runge-Kutta method. In the following section, we will briefly introduce Runge-Kutta methods and some of their properties, thus deriving the notation above. The probabilistic method we considered in this work, first introduced in \cite{CGS16}, considers independent identically distributed (i.i.d.) random variables $\xi_k$, which will explicitly depend on the discretization step $h$, and add them at each step of the numerical integration, i.e., 
\begin{equation}\label{eq:probMethod}
\begin{aligned}
	U_{k+1} &= \Psi(U_k) + \xi_k(h), \\
	U_0 &= u_0.
\end{aligned}
\end{equation}
Hence, the probabilistic solution will depend on the realizations of the random variables $\xi_k$, thus providing a family of numerical solutions instead of a punctual solution as in deterministic numerical analysis theory. \\
In the following section, we will briefly present Runge-Kutta methods we considered further in this work, summarizing some properties defining these methods. Then, we will present a motivating example for the probabilistic solver we introduced above. Finally, we will present some properties of the method together with numerical experiments empirically confirming our analytical investigations. 

\input{deterministicmethods}
\input{ProbabilisticMotivation}

\subsection{Method properties}
The properties of the probabilistic method introduced in \eqref{eq:probMethod} have been partially investigated in \cite{CGS16}, where a result on strong and weak convergence of the numerical solution is derived. In the following, we will briefly introduce this analysis and show a numerical example that confirms the theoretical results. \\
A property which conversely has not been investigated is the convergence of a Monte Carlo approximation of the numerical solution. Let us consider the distribution $\prior_h$ given by the probabilistic solver, and the distribution $\delta_u$ representing the Dirac delta concentrated on the true solution. In \cite{CGS16}, and as we will show in the following, the convergence of $\prior_h$ to $\delta_u$ is theoretically described. As pointed out in \cite{KeH16}, we can only access a distribution $\prior_h^M$ through $M$ repeated samples of the probabilistic solution, which will approximate the distribution $\prior_h$. As it is schematically described in \cite{KeH16}, the two following convergences have to be analyzed
\begin{equation}
	\prior_h^M \totext{h\to 0, M\to \infty} \prior_h \totext{h\to 0} \delta_u.
\end{equation}
In \cite{CGS16}, only the second convergence is treated. In the following sections, we will show that in case the initial condition is deterministic, the convergence of $\prior_h^M$ to $\prior_h$ is independent of the number of samples $M$, and therefore the computational cost due to the approximation of $\prior_h$ through repeated sampling is negligible. \\
Finally, we will discuss the case where this favorable convergence property is not valid together with an application of multi-level techniques, and we will conclude with a brief analysis of the stability of this numerical method. 

\input{strongconvergence}
\input{weakconvergence}
\input{MonteCarlo}
\input{MLMC}
\input{Stability}




